---
title: "NYPD_Project"
author: "Anonymous"
date: "6/28/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## NYPD Shooting Incidents Project
This project intends to investigate data on shooting incidents in the area of New York City, U.S.A. The dataset is publicly available on the website of the City of New York. It lists every shooting incident that occurred in NYC from 2006 through the end of 2020. According to the website "each record represents a shooting incident in NYC and includes information about the event, the location and time of occurrence [and] information related to suspect and victim demographics is also included".

### Reading In Data
I start by reading in the data from the csv file that is provided by the City of New York at https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv. I need to load the **tidyverse** first.

```{r get_ny_data}
library(tidyverse)
url_in <- "https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv"
NYPD_data <- read_csv(url_in)
str(NYPD_data)
summary(NYPD_data)
```
### Remove Unnecessary Columns
Now I get rid of the *COORD* columns as well as the *Latitude* and *Longitude* columns and the columns *PRECINCT*, *JURISDICTION_CODE* and *LOCATION_DESC* since I do not need them in my further analysis. Since it might be interesting to analyse at what time of the day the most incidents occur, I will keep the *OCCUR_TIME* column.

```{r rm_cols}
NYPD_data <- NYPD_data %>% select(-c('PRECINCT', 'JURISDICTION_CODE', 'LOCATION_DESC', 'X_COORD_CD', 'Y_COORD_CD', 'Latitude', 'Longitude', 'Lon_Lat'))
summary(NYPD_data)
```

Looking at the data we see that most columns are of the character type. However it makes sense to use factors, and a date type for the *OCCUR_DATE*.

### Date Type
Let us start with the date and time. We will take advantage of **lubridate**.

```{r adj_data}
library(lubridate)
NYPD_data <- NYPD_data %>% mutate(OCCUR_DATE = mdy(OCCUR_DATE))
NYPD_data <- NYPD_data %>% mutate(OCCUR_TIME = hms(OCCUR_TIME))
summary(NYPD_data)
```
### Create Factors

First, the boroughs (*BORO*). There are 5 entries. I will create a factor from the entries. For *STATISTICAL_MURDER_FLAG* there are two possible values, *true* (murder) or *false* (no murder). I will create a factor. If we look at the *PERP_AGE_GROUP*, there are obviously missing values and strange values (1020, 940 and 224). I will look at this later on and just create a factor for now. For *PERP_SEX* there are also missing values. I will create a factor for now. I will also create factors for the remaining columns.

```{r unique_values_boro}
unique(NYPD_data$BORO)
NYPD_data$BORO <- as.factor(NYPD_data$BORO)

unique(NYPD_data$STATISTICAL_MURDER_FLAG)
NYPD_data$STATISTICAL_MURDER_FLAG <- as.factor(NYPD_data$STATISTICAL_MURDER_FLAG)

unique(NYPD_data$PERP_AGE_GROUP)
NYPD_data$PERP_AGE_GROUP <- as.factor(NYPD_data$PERP_AGE_GROUP)

unique(NYPD_data$PERP_SEX)
NYPD_data$PERP_SEX <- as.factor(NYPD_data$PERP_SEX)

unique(NYPD_data$PERP_RACE)
NYPD_data$PERP_RACE <- as.factor(NYPD_data$PERP_RACE)

unique(NYPD_data$VIC_AGE_GROUP)
NYPD_data$VIC_AGE_GROUP <- as.factor(NYPD_data$VIC_AGE_GROUP)

unique(NYPD_data$VIC_SEX)
NYPD_data$VIC_SEX <- as.factor(NYPD_data$VIC_SEX)

unique(NYPD_data$VIC_RACE)
NYPD_data$VIC_RACE <- as.factor(NYPD_data$VIC_RACE)

summary(NYPD_data)
```
There is only missing data for the perpetrator columns while the victim data is complete.

### Visualization & Analysis
For my visualizations and analyses I want to focus on the **murder cases**. Therefore I will create a new dataset that contains all the murder cases.

```{r muder_cases_dataset}
NYC_murders <- NYPD_data %>% filter(STATISTICAL_MURDER_FLAG==T)
summary(NYC_murders)
```

Now I will visualize the number of murders per borough.

```{r murders_by_borough_visual}
ggplot(NYC_murders) +
  geom_bar(aes(y = BORO, fill = BORO)) +
  scale_x_continuous(breaks=c(seq(0, 2000, by=250)), labels=c(seq(0, 2000, by=250))) +
  labs(title = "Total Murder Cases per Borough - 2006-2020", x = "Murder Cases", y = "NYC Borough")
```

Most murders happened in Brooklyn. Now, it would be a good idea to incorporate the population into the dataset and calculate the incidents per 100,000 people. However, this would go too far for now in this project since we would need the data for all the relevant years.

It will be interesting to see how many murder cases there were per year.

```{r murders_by_year_visual}
ggplot(NYC_murders, aes(x=year(OCCUR_DATE))) +
  geom_bar(fill = "coral1") +
  scale_x_continuous(breaks=c(2006:2020), labels=c(2006:2020)) +
  labs(title = "Total Murder Cases per Year - 2006-2020", x = "Year", y = "Murder Cases")
```

This plot shows that the number of murders declined from 2006-2019 but then in 2020 it almost doubled compared to 2017, 2018 and 2019, respectively.

Now I want to visualize the number of murders by month to see whether there are months that are more dangerous for murders to occur than others.

```{r murders_by_month_visual}
ggplot(NYC_murders, aes(x=month(OCCUR_DATE))) +
  geom_bar(fill = "coral1") +
  scale_x_discrete(limits=month.abb) +
  labs(title = "Total Murder Cases per Month - 2006-2020", x = "Month", y = "Murder Cases")
```

It is obvious that the summer months seem to be the most dangerous months. On the other hand, the beginning of the year seems to be less dangerous than the end of the year when it comes to murders.

Another interesting question would be: "How did properties like age, sex and race change (or not change) over time?" And many more.

### Model
I will use binary logistic regression to answer the question if there are properties that make it more likely for a shooting incident to be a murder case. For this model I will take *STATISTICAL_MURDER_FLAG* as the independent variable (*TRUE* means murder) and *PERP_SEX*, *PERP_RACE* and *BORO* as independent variables.

```{r model_murders_regression}
murder_reg_mod <- glm(STATISTICAL_MURDER_FLAG ~ PERP_SEX + PERP_RACE + BORO, family=binomial, data=NYPD_data)
summary(murder_reg_mod)
```

A *Pr* value of less than 0.05 means that the independent variable has statistically significant influence on the dependent variable. The summary shows that male perpetrators have a significant higher probability of being involved in a murder case when a shooting incident appears. We also see that in the borough of Brooklyn the likelihood of a shooting incident to be a murder case is significant higher. Interestingly, there are no signs that the perpetrator's race has any influence on a shooting incident being a murder case.

### Bias
The first issue that comes to my mind when thinking about bias in this dataset is regarding race. From just looking at the data one might get the impression that race could be the reason for the number of shooting incidents while in fact it depends mostly on other social and economic factors. Another source for bias could be the sex of the perpetrator or the victim. And missing values are a problem in this dataset since sex, age group and race information is missing for over 8400 perpetrators.

### Conclusion
This project has given an overview about how to begin data analysis. From reading in the data over tidying it up to visualizing some simple but interesting facts. This will establish a good foundation for future projects.